---
title: "IF_NAFLD_cellchat_v1"
output: html_document
date: "2025-09-12"
editor_options: 
  chunk_output_type: console
---

# Load libraries
```{r}
library(Seurat)
library(ggplot2)
library(SingleCellExperiment)
library(Scillus)
library(dplyr)
library(tidyr)
library(Cairo)
library(presto)
library(ComplexHeatmap)
library(grid)
library(CellChat)
```

# Organize seurat identities
```{r}
Idents(liver_Chow) <- liver_Chow$CellType
liver_Chow <- RenameIdents(liver_Chow, 
                      "B" = "B cells",
                      "DCs" = "Dendritic cells",
                      "pDCs" = "Plasma dendritic cells",
                      "T" = "T cells",
                      "NK" = "NK cells",
                      "Monocyte/Monocyte derived macrophage" = "Monocytes/Macrophages")

levels(liver_Chow) <- c('Hepatocytes', 
                   'Hepatic progenitor cells', 
                   'Stellate cells', 
                   'Myofibroblasts',
                   'Neutrophils',
                   'Monocytes/Macrophages',
                   'Kupffer cells',
                   'Dendritic cells', 
                   'Plasma dendritic cells',
                   'NK cells',
                   'B cells',
                   'T cells',
                   'Endothelial cells')
liver_Chow$CellType <- Idents(liver_Chow)
```

# Subset Chow and 15 week data containing cells of interest
```{r}
liver_chow_coi <- subset(
  liver_Chow,
  idents = c(
    "Hepatocytes",
    "Stellate cells",
    "Myofibroblasts",
    "Neutrophils",
    "Monocytes/Macrophages",
    "Kupffer cells"
  )
)

liver_15weeks_coi <- subset(
  liver_15weeks,
  idents = c(
    "Hepatocytes",
    "Stellate cells",
    "Myofibroblasts",
    "Neutrophils",
    "Monocytes/Macrophages",
    "Kupffer cells"
  )
)
```


# Create cell chat object from Seurat
```{r}
cc_chow <- createCellChat(object = liver_Chow,
                           meta = liver_Chow@meta.data,
                           group.by = "CellType")
cc_15w <- createCellChat(object = liver_15weeks,
                           meta = liver_15weeks@meta.data,
                           group.by = "CellType")
```

# Set working database
```{r}
## Set the ligand-receptor interaction database
CellChatDB <- CellChatDB.mouse
# use CellChatDB.human  if running on human data
showDatabaseCategory(CellChatDB)
```

# Examine database
```{r}
# Show the structure of the database
dplyr::glimpse(CellChatDB$interaction)

# Save this data: Txt file
write.table(
  CellChatDB$interaction,
  file = "CellChatDB.txt",
  sep = "\t",
  row.names = TRUE,
  col.names = NA
)
```

# Define database to use
```{r}
#use all CellChatDB except for "Non-protein Signaling" for cell-cell communication analysis
CellChatDB.use <- subsetDB(CellChatDB)
showDatabaseCategory(CellChatDB.use)
write.table(
  CellChatDB.use$interaction,
  file = "CellChatDB_no_nonprotein.txt",
  sep = "\t",
  row.names = TRUE,
  col.names = NA
)
```

# Set used database in cellchat object
```{r}
cc_chow@DB <- CellChatDB.use
cc_15w@DB <- CellChatDB.use
```

# Confirm database used in object of interest
```{r}
showDatabaseCategory(cc_15w@DB)
```

## Preprocessing the expression data for cell-cell communication analysis

# subset the expression data of signaling genes for saving computation cost
```{r}
cc_chow <- subsetData(cc_chow)
cc_15w <- subsetData(cc_15w)
```

# Identify over expressed genes/interactions
```{r}
cc_chow <- identifyOverExpressedGenes(cc_chow)
cc_chow <- identifyOverExpressedInteractions(cc_chow)

cc_15w <- identifyOverExpressedGenes(cc_15w)
cc_15w <- identifyOverExpressedInteractions(cc_15w)
```

# Compute the communication probability and infer cellular communication network
```{r}
cc_chow <- computeCommunProb(cc_chow, type = "triMean")
cc_15w <- computeCommunProb(cc_15w, type = "triMean")
```

## Infer the cell-cell communication at a signaling pathway level
```{r}
cc_chow <- computeCommunProbPathway(cc_chow)
cc_15w <- computeCommunProbPathway(cc_15w)
```

# Calculate the aggregated cell-cell communication network
```{r}
cc_chow <- aggregateNet(cc_chow)
cc_15w <- aggregateNet(cc_15w)
```

# Calculate compute centrality
```{r}
cc_chow <- netAnalysis_computeCentrality(cc_chow, slot.name = "netP")
cc_15w <- netAnalysis_computeCentrality(cc_15w, slot.name = "netP")
```

# Merge datasets of interest
```{r}
list_allcells <-
  list(Chow = cc_chow, HFD_15wk = cc_15w)
cc_chow_15w <-
  mergeCellChat(list_allcells, add.names = names(list_allcells))
```