---
title: "IF_NAFLD_scRNAseq_v1"
output: html_document
date: "2025-09-29"
---

# Load libraries
```{r}
library(Seurat)
library(ggplot2)
library(SingleCellExperiment)
library(Scillus)
library(dplyr)
library(tidyr)
library(Cairo)
library(presto)
library(ComplexHeatmap)
library(grid)
library(CellChat)
```

# Read in the counts file and make it a sparse matrix format
```{r}
liver_counts <- read.table(file = "/Volumes/Samsung_T5/20230518_MacBook_public_snRNAseq/8_QiSu_iScience_liver_NAFLD_mouse/1_rawdata/output/GSE166504_cell_raw_counts.20220204.txt", 
                     sep = '\t', 
                     header = T) # if .csv, sep = ","

liver_counts <- read.table(file = "counts_matrix.txt", 
                     sep = '\t', 
                     header = T,
                     row.names = liver_counts$X)

liver_counts$X <- NULL

liver_counts_mx <- as.matrix(liver_counts)
```

# Read meta.data
```{r}
liver_meta.data <- read.table("/Volumes/Samsung_T5/20230518_MacBook_public_snRNAseq/8_QiSu_iScience_liver_NAFLD_mouse/1_rawdata/output/GSE166504_cell_metadata.20220204.tsv", 
                              sep = '\t', 
                              header = T)

View(liver_meta.data)

sampleinfo <- liver_meta.data$FileName
spec_sampleinfo <- str_split_fixed(sampleinfo, "_", 4)

liver_meta.data$input <- spec_sampleinfo[,1]
liver_meta.data$food <- spec_sampleinfo[,2]
liver_meta.data$animal <- spec_sampleinfo[,3]
liver_meta.data$capture <- spec_sampleinfo[,4]

View(liver_meta.data)

row.names(liver_meta.data) <- make.unique(liver_meta.data$CellID)
liver_meta.data$CellID <- NULL
liver_meta.data$FileName <- NULL

View(liver_meta.data)
```

# Create a seurat object using expression matrix and meta.data
```{r}
liver <- CreateSeuratObject(counts = liver_counts_mx,
                           assay = "RNA",
                           meta.data = liver_meta.data,
                           min.cells = 3,
                           project = "liver")
str(liver)

id <- colnames(liver_counts_mx)

table(nchar(liver_meta.data$CellID))

id_barcode <- substr(id, nchar(id)-16+1, nchar(id))
table(nchar(id_barcode))

colnames(liver_counts_mx) <- make.unique(id_barcode)

liver <- CreateSeuratObject(counts = liver_counts_mx,
                           assay = "RNA",
                           meta.data = liver_meta.data,
                           project = "liver")
```

# Quality Control : take a look at the paper to see parameters used
"cells with fewer than 300 detected genes, more than 60,000 detected UMI, and a very high (>0.8) mitochondrial genome transcript ratio were filtered.
```{r}
liver$percent.mito <- PercentageFeatureSet(liver, pattern = "^mt-")
# no genes. Let's double check data

VlnPlot(liver, c("nFeature_RNA", "nCount_RNA"))
# Load data
```{r}
load("liver_seuratonly_20240102.rds")
```

```{r}
# Normalize data
```{r}
liver <- NormalizeData(liver)
```

# Find Variable Features
paper : "The top 2,000 variable genes(HVGs) were identified using the function FindVariableFeatures"
```{r}
liver <- FindVariableFeatures(liver, nfeatures = 2000)
```

# Scale Data
```{r}
liver <-
  ScaleData(liver, vars.to.regress = c("food", "capture", "animal"))
```

# Perform PCA & Run UMAP
```{r}
liver <- RunPCA(liver)
liver <- RunUMAP(liver, dims = 1:30)
```



